---
title: "SDS 315 HW 2"
author: "Suryabh Singh"
date: "2025-01-28"
output:
html_document:
df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Your Name and GitHub Link

**Name:** Your Full Name\
**UT EID:** Your UT EID\
**GitHub Repository:** [GitHub Link](https://github.com/yourrepo)

## Analysis and Plots

### Load Data

```{r}
library(ggplot2)
library(dplyr)

# Load the profs.csv dataset
profs <- read.csv("profs.csv")
```

### Part A: Histogram of Course Evaluation Scores

```{r}
ggplot(profs, aes(x = eval)) + 
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") + 
  labs(
    title = "Distribution of Course Evaluation Scores",
    x = "Evaluation Score",
    y = "Frequency"
  ) + 
  theme_minimal()
```

**Caption:** The histogram shows the overall distribution of course evaluation scores. Most scores are clustered around 4, suggesting generally positive evaluations.

### Part B: Boxplots by Native English Speaker

```{r}
ggplot(profs, aes(x = factor(native, labels = c("Non-Native", "Native")), y = eval)) + 
  geom_boxplot(fill = c("#FF9999", "#99CCFF")) + 
  labs(
    title = "Course Evaluation Scores by Native English Speaker Status",
    x = "Native English Speaker",
    y = "Evaluation Score"
  ) + 
  theme_minimal()
```

**Caption:** Side-by-side boxplots display the distribution of course evaluation scores based on whether professors are native English speakers. Native speakers tend to have slightly higher median evaluation scores.

### Part C: Faceted Histogram by Gender

```{r}
ggplot(profs, aes(x = eval)) + 
  geom_histogram(binwidth = 0.2, fill = "steelblue", color = "black") + 
  facet_wrap(~gender, nrow = 2, labeller = as_labeller(c("female" = "Female", "male" = "Male"))) + 
  labs(
    title = "Distribution of Course Evaluation Scores by Gender",
    x = "Evaluation Score",
    y = "Frequency"
  ) + 
  theme_minimal()
```

**Caption:** The faceted histograms compare evaluation score distributions between male and female instructors. Both distributions are centered around 4, with slight variations.

### Part D: Scatterplot of Attractiveness vs. Evaluation

```{r}
ggplot(profs, aes(x = beauty, y = eval)) + 
  geom_point(alpha = 0.6, color = "darkorange") + 
  geom_smooth(method = "lm", se = TRUE, color = "blue") + 
  labs(
    title = "Physical Attractiveness vs. Course Evaluation Scores",
    x = "Physical Attractiveness",
    y = "Evaluation Score"
  ) + 
  theme_minimal()
```

**Caption:** The scatterplot shows a potential positive association between professors' physical attractiveness and their course evaluation scores. The trendline indicates that higher attractiveness may correlate with better evaluations.
